---
title: "Univariate Analysis"
output: html_document
date: "2024-09-30"
---
```{r}
library(moments)
```

```{r}
# Reading in the data
Sale_Price <- read.csv("Sales_data.csv",header=T)

```

# Summary Statistics
```{R}
summary(Sale_Price)
```

```{R}

# Calculate and print summary statistics for unit_sale_price_gbp
mean_unit_sale_price_gbp <- mean(Sale_Price$unit_sale_price_gbp)
median_unit_sale_price_gbp <- median(Sale_Price$unit_sale_price_gbp)
sd_unit_sale_price_gbp <- sd(Sale_Price$unit_sale_price_gbp)
var_unit_sale_price_gbp <- var(Sale_Price$unit_sale_price_gbp)
skewness_unit_sale_price_gbp <- skewness(Sale_Price$unit_sale_price_gbp)
kurtosis_unit_sale_price_gbp <- kurtosis(Sale_Price$unit_sale_price_gbp)

# Print the results
print(paste0("Mean: ", mean_unit_sale_price_gbp))
print(paste0("Median: ", median_unit_sale_price_gbp))
print(paste0("Standard Deviation: ", sd_unit_sale_price_gbp))
print(paste0("Variance: ", var_unit_sale_price_gbp))
print(paste0("Skewness: ", skewness_unit_sale_price_gbp))
print(paste0("Kurtosis: ", kurtosis_unit_sale_price_gbp))
```

# Analysis of results
Adressing the skewness of the data. The mean of the data is larger than the median, as wel las a skewness value of 1.38 suggesting that the data is *right skewed*

```{R}
# Histogram
hist(Sale_Price$unit_sale_price_gbp,main = "Histogram of Sale Price" ,xlab = "Sales Price", ylab = "Frequency", col = "lightblue")
```

This histogram has a long tail towards the right (right skew). This can be stastically represented by Kurtosis of 5.21 which indicates the heavy tail.

```{R}
# Density Plot
plot(density(Sale_Price$unit_sale_price_gbp), main = "Density Plot of Sales Prices", xlim = c(0,70))

```

# Testing for Normality
The normality of data can affect the choice of stastical tests and models created. The shape of the data suggests that it does not follow a normal distribution but there are additional tests that we can use. The example I've used here is a *Shapiro-Wilk* test
```{R}
# Shapiro Test
shapiro.test(Sale_Price$unit_sale_price_gbp)

```

This test gave a very small p value. Meaning we can say with confidence that the data is not normally distributed. This can be visualised in a Q-Q plot. The closer the points are to the normal line, the more likely the data follows a normal distribution 
```{R}
# Q-Q Plot
qqnorm(Sale_Price$unit_sale_price_gbp)
qqline(Sale_Price$unit_sale_price_gbp)
```

In this diagram, the data deviated from the line in the 2nd and 3rd quartiles, again suggesting it does not follow a normal distribution. It also helps us to confirm the right skew of the data 

# Investigating outliers
We know from the shape of the histogram and the Kurtosis value that there are some large values that are skewing the data. These can be visualised by a boxplot 
```{R}
# Boxplot
boxplot(Sale_Price$unit_sale_price_gbp,main = "Box Plot of Sales Prices", col = "lightblue")
```

This boxplot shows that there are values that are outside limits to be considered an outlier and and will skew the data.

# Conclusions 
This dataset has a wide variance, numbers are spread far out and there are extremes at the larger end that are causing right skew

The mean of this data will not be ideal to use as there are distruptions to the *central tendancy of the data*. In this case it will be most useful to use the **median** instead of the mean. As it is less affected by the skew of the data and will be more representative of an average unit price

## Some considerations
With larger datasets median may not be so effective 
1. Computational power used to calculate the median is higher
2. Medians are less useful with lots of zero values (median number of pregnancies per copulation is zero)
3. There are some data types where median is as useful. E.g. In sales, removing the tails might underestimate your
